---
title: "Binomial test"
output: html_notebook
---

## Binomial distribution
### Problem 1
A fair coin is tossed 7 times. Find probability that

- we obtained 7 heads;
- we obtained 0 heads;
- we obtained 1 head;
- we obtained 6 heads;
- we obtained 6 heads or more;
- we obtained 1 head or less;
- we obtained 6 heads or less;
- we obtained 5 heads or less.

Solve this problem using R as calculator (without any functions)

### Problem 2
There is a box with two identical red balls and one blue ball. The box is shuffled, then one ball is picked from the box randomly. Its color is recorded, then the ball is returned to the box. The process is repeated three times. Find probability that

- We picked blue balls three times.
- We picked red balls three times.
- We picked two red balls and one blue ball.
- We picked two twe blue and one red balls.
- We picked two or more blue balls.
- We picked two or less blue balls.

Solve this problem using R as calculator (without any functions)

### Problem 3
Discuss functions `rbinom`, `dbinom`, `pbinom`. Visualize the binomial distribution for different valeus of $p$ and $n$. (Use `rbinom` and show that the result of simulation is similar to the probabilities obtained by `dbinom`.) Use appropriate functions to solve some points from Problem 1 and Problem 2. 

### Problem 4
Consider the magician who claims he can predict future. Our null hypothesis is that he actually cannot predict future and therefore probability to predict correctly the result of one coin tossing is 1/2. An alternative, according to magician's claim, is that this probability is larger than 1/2. (It's not 1 because magician can make mistakes from time to time (future-telling is hard after all). Assume that we tossed a coin 4 times and the magician guessed all 4 results. Find p-value. Would you reject null hypothesis with 5% significance level? Would you claim the magician indeed has magical abilities?

### Problem 5
What if the magician guessed 5 times out of 5?

### Problem 6
What if the magician guessed 9 times out of 11?

### Problem 7
In a different experiment, the magician tries to guess the result of the ball picking experiment described in Problem 2. What is his best strategy provided he doesn't have magical abilities? Assume that he guessed 7 times out of 7 using the best strategy. State the null hypothesis and find p-value. Would you claim that he has magical abilities? Use 5% significance level.

### Problem 8
Use function `binom.test` to solve problems 4 — 7.

### Problem 9

The dataset [“The last words in verses”](https://raw.githubusercontent.com/LingData2019/LingData2020/master/data/poetry_last_in_lines.csv) by Olga Lyashevskaya is based on texts written in the 1820s and 1920s (Corpus of Russian Poetry of the Russian National Corpus). Only one line per author is collected to keep observations as independent from each other as possible. 

Variables:  

- `Decade` — decade of creation: 1820s, 1920s.  
- `RhymedNwords` — the number of words in the rhyming position (usually one word, but there are two words in cases such as _вина бы_ 'I would like to get) wine' (which is rhymed with _жабы_ 'toad', see http://russian-poetry.ru/Poem.php?PoemId=18261)).  
- `RhymedNsyl` — the number of syllables in the rhyming position.  
- `UPoS` — part of speech of the last word. 
- `LineText` — a sampled verse.  
- `Author` — the author of the text.  

Can we decide that in verses written in 1920s, verbs in the rhyming position are used less often than expected for verbs in general? 

Let's assume that the probability for verbs to be used in any position ('in general') is 17% (according to http://www.ruscorpora.ru/new/corpora-stat.html).

#### 9.1 State hypothesis

What is your null hypothesis $H_0$ and what is the alternative hypothesis $H_1$?

#### 9.2. 

Read the dataset [“The last words in verses”](https://raw.githubusercontent.com/LingData2019/LingData2020/master/data/poetry_last_in_lines.csv). Filter out the relevant observations from 1920s, calculate the number of verbs observed in the sample, and the sample size. 

```{r}

```

#### 9.3

Use an exact binomial test to calculate p-value. 
```{r}
```

#### 9.4 Interpret results

Give your interpretation of obtained p-value. Answer the initial question: Can we decide that in verses written in 1920s, verbs are used in the rhyming position less often than expected?
