---
title: "HW 3. Binomial test and t-test"
header-includes:
  - \usepackage{fontspec}
  - \setmainfont{CMU Serif}
  - \setsansfont{CMU Sans Serif}
  - \setmonofont{CMU Typewriter Text}
  - \newcommand{\tightlist}{%
  - \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
output:
  pdf_document:
    latex_engine: lualatex
  html_document: default
---

_This homework is based on the homeworks given to students of the same course in 2019 teached by Olga Lyashevskaya, George Moroz, Alla Tambovtseva and Ilya Schurov._


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
```

## Your name:

## 1. Position of verbs in verses
### Description of dataset
The dataset [“The last words in verses”](https://raw.githubusercontent.com/LingData2019/LingData/master/data/poetry_last_in_lines.csv) contains a sample of lines taken from the RNC Corpus of Russian Poetry. Actually, there are two samples comprising the texts written in the 1820s and 1920s. We took only one line per author to keep our observations as independent as possible.

Variables:

- Decade — decade of creation: 1820s, 1920s.
- RhymedNwords — the number of words in the rhyming position (usually one word, but there are two words in cases such as _вина бы_ 'I would like to get) wine' (which is rhymed with _жабы_ 'toad', see http://russian-poetry.ru/Poem.php?PoemId=18261)).
- RhymedNsyl — the number of syllables in the rhyming position.
- UPoS — part of speech of the last word.
- LineText — a sampled verse.
- Author — the author of the text.

### Research question
We are interested in the following question:
can we decide that in verses written in 1820s, verbs are used in the rhyming position more often or less often than expected for verbs in general?

### 1.1 General expectations

To calculate the probability to come across a verb in written Russian texts (general expectations), use the [frequency dictionary of the Russian National Corpus](https://raw.githubusercontent.com/LingData2019/LingData/master/data/freq_rnc_ranked.csv) (Lyashevskaya, Sharoff 2009).

#### 1.1.1 Read the file
First of all, we have to read the RNC frequency dictionary data.

We will use some functions from `tidyverse` library, so do not forget to put line `library(tidyverse)` in the beginning of your code

Note that the file is tab-separated, so we have to use `read_tsv` instead of `reac_csv` function in order to read it.

You probably also need to activate UTF-8 locale using command
```
Sys.setlocale(locale='UTF-8')
```
to see cyrillic letters in `Lemma` column (but it is not necessary for this exercise).

Read [frequency dictionary](https://raw.githubusercontent.com/LingData2019/LingData/master/data/freq_rnc_ranked.csv) file and save the result as variable `freq_rnc`.

```{r read-rnc-dataset}
# YOUR CODE HERE
```

#### 1.1.2 Investigate dataset
Use `head` command to show the beginning of the table stored in variable `freq_rnc`.

```{r}
head(freq_rnc)
```

Note that verbs are coded as `'v'` in the `PoS` field, and their frequency is shown in the `Freq(ipm)` field (relative frequency, number of items per million words in the corpus).

#### 1.1.3. Find the probability

Calculate the probability to see verbs dividing the sum of their frequency by the sum of frequency of all words in the dictionary. Note that due to presence of brackets in the name of column `Freq(itm)`, to access this column, you have to put its name into backticks, i.e.:

```{r}
all_freqs <- freq_rnc$`Freq(ipm)` # access column Freq(ipm)
all_freqs[1:10] # show first 10 elements
```

```{r rnc-p-verbs}
# YOUR CODE HERE
```

### 1.2 State hypothesis

Assume that every time the author decides which word to use as a last word of a verse, they toss a coin. If they get head, they use verb, otherwise they use word of different part of speech. Denote the probability to get head (i.e. use verb) by $p$. Recall that we are interested in the following question: can we decide that in verses written in 1820s, verbs are used in the rhyming position more often or less often than expected for verbs in general?

State null hypothesis $H_0$ and alternative $H_1$ in terms of $p$.

#### Your answer

### 1.3 Analyse data
#### 1.3.1. Read the dataset
Read the dataset [“The last words in verses”](https://raw.githubusercontent.com/LingData2019/LingData/master/data/poetry_last_in_lines.csv) and put it variable `lvw`. Filter out the relevant observations from 1820s, calculate the number of verbs observed in the sample, and the sample size.

Hint. You can use function `table` to calculate how many times each value occurs in a vector of factors.

```{r read-poetry-dataset}
# YOUR CODE HERE
```

### 1.3.2 Do statistical testing

Test stated hypothesis and find p-value. Use `binom.test` that will calculate p-value for binomial test (note that it uses two-sided alternative by default). You have to put actual number of successes (i.e. _heads_), number of trials (i.e. coin-tossings) and expected probability of success that is used in null hypothesis.


```{r p-value}
# YOUR CODE HERE
```

### 1.4 Interpret results

Give your interpretation of obtained p-value. Would you reject null hypothesis? Answer the initial question: can we decide that in verses written in 1820s, verbs are used in the rhyming posision more often or less often than expected?

#### Your answer

## 2. Frequent words, their acoustic duration and co-articulation effects

Many studies report shorter acoustic durations, more co-articulation and reduced articulatory targets for frequent words. The study of [Fabian Tomaschek et al.](https://www.semanticscholar.org/paper/Practice-makes-perfect-%3A-The-consequences-of-for-Tomaschek-Tucker/1e0dbc3787a6da84ffd4c3cae62f1340e4267694) (2018) investigates a factor ignored in discussions on the relation between frequency and phonetic detail, namely, that motor skills improve with experience.

For this research people were asked to read texts with target German verbs aloud and then the duration of their speech was recorded. Participants had to speak in different conditions, slow and fast. In other words, they were asked to speak slowly/fast or the setting for speaking slowly/fast was created implicitly (so speakers did not understand that).

In this homework you are suggested to compare word duration and text segment duration for fast an slow speaking conditions. On the one hand, it is logical to suppose even without testing that duration in fast speaking condition should be shorter. On the other hand, before doing a more substantial research it might be helpful to check whether this intuitive suggestion holds, i.e. to make sure that the conditions of the experiment were thoroughly maintained (for example, researchers did not swap conditions and recorded results correctly).

**Variables of interest:**

* `LogDurationW` - log-transformed word duration (i.e. logarithms of word duration).
* `LogDurationA` - log-transformed segment duration.
* `Cond` - condition (slow, fast).

### 2.0 Data loading

Load data ([link](http://math-info.hse.ru/f/2018-19/ling-data/dur_word_frequency.csv)), save it to variable `dur_word_freq` and print the beginning of this dataset with `head`. (Note that this is csv-file, not tab-separated, so you have to use function `read_csv`.)

```{r load-data}
# YOUR CODE HERE
```


For brevity, below we will refer to variables `LogDurationW` and `LogDurationA` as "word duration" and "segment duration" correspondingly despite the fact that they are actually logarithms of the durations.

### 2.1 Word duration and segment duration

Draw histograms for word duration and segment duration values.

```{r hist}
# YOUR CODE HERE
```


### 1.2 Segment duration in slow and fast condition

Run the following code:

```{r duration}
boxplot(LogDurationA ~ Cond, data=dur_word_freq)
```

The result is so-called _box and whisker plot_ for variable `LogDurationA` groupped by variable `Cond`. You can read about meaning of the elements of box plot in [Wikipedia article](https://en.wikipedia.org/wiki/Box_plot) (also, [in Russian](https://ru.wikipedia.org/wiki/Ящик_с_усами)). What can you say about difference between values of `LogDurationA` that correspond to fast and slow conditions? Is it reasonable to expect that segment durations are shorter for fast speaking condition than for slow speaking condition? Can the graph you plotted confirm this? What kind of assertions can you make from the graph? E.g. can you assert something like "sample/population mean/median of segment duration in fast speaking condition is shorter/longer than in slow speaking condition"?

#### Your answer

### 2.3. Word duration in slow and fast condition
Repeat 1.2 using word duration instead of segment duration. Run appropriate code and interpret the resulting figure.
```{r}
# YOUR CODE HERE
```

#### Your answer (interpretation of a graph)

### 3.1 Student's t-test

Now using Student's t-test we want to decide whether the difference between

(a) word duration in fast condition and word duration in slow condition,

(b) segment duration in fast condition and segment duration in slow condition

is statistically significant. In other words, we want to check is it true that these durations differ not only in the samples, but also in the populations.

#### 3.1.1 Hypothesis

First of all, state the null hypothesis and the alternative you consider (both for cases (a) and (b) above). Justify your choice of alternative hypothesis.

##### Your answer

#### 3.1.2 Application of test

Apply `t.test` to check the hypothesis (both for cases (a) and (b) above).

```{r t-tests}
# YOUR CODE HERE
```

#### 3.1.3 Interpretation

Interpret results of the t-test performed. Report p-values obtained. Can you confirm that there is a difference between word duration in fast condition and word duration in slow condition in the population? The same question for the segment duration.

##### Your answer
